---
title: "HW1_wliu3"
author: "Wei Liu"
date: "8/24/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Problem 1
Finished in the RStudio Cloud

# Problem 2
### Part A
My 3 desired learning objectives are as follows:

\begin{enumerate}
 \item have a better understanding of tools that data scientists work with
 \item be able to use these tools to explore the process of data analysis
 \item be able to communicate with my collaborators efficiently using these tools
\end{enumerate}

### Part B

$y_1$:\ uniform distribution\
$y_2$:\ exponential distribution\
$y_3$:\ pareto distribution

\begin{eqnarray}
y_1 &=& \frac{1}{2},~~~ 2\leq x\leq 4\\
y_2 &=& \frac{1}{2}exp\left ( \frac{-x}{2} \right ),~~~ 0\leq x\leq \infty\\
y_3 &=& \frac{2}{x^{3}},~~~ 1\leq x\leq \infty
\end{eqnarray}

# Problem 3

Summary of the steps in performing Reproducible Research

\begin{enumerate}
 \item For Every Result, Keep Track of How It Was Produced
Comment: It's a good way to comment the meaning or purpose of important code to make it clear and understandable.
 \item Avoid Manual Data Manipulation Steps
Comment:For problem 4, there is no such problem since an internal R dataset is used.
 \item Archive the Exact Versions of All External Programs Used
Comment: Good suggestion since there might be different functions in different version.
 \item Version Control All Custom Scripts
Comment: Github is an efficient example tool to keep track all changes and scripts we made, so we can check back anytime and anywhere.
 \item Record All Intermediate Results, When Possible in Standardized Formats
Comment: Intermediate results are important because they might provide insights for our final results and they allow parts of process to run firstly!
 \item For Analyses That Include Randomness, Note Underlying Random Seeds
Comment: Nice point! Record the seed number to make sure get reproducible results.
 \item Always Store Raw Data behind Plots
Comment: Good habit to store raw data behind plots in case further investigation.
 \item Generate Hierarchical Analysis Output, Allowing Layers of Increasing Detail to Be Inspected
Comment: Put as much details since all of them maybe useful in future.
 \item Connect Textual Statements to Underlying Results
Comment: Textual statements is a good way to communicate with others.
 \item Provide Public Access to Scripts, Runs, and Results
Comment: Set the access as public such as in Github to make the resources accessible to readers.
\end{enumerate}

# Problem 4
**Scatter plot**
```{r, fig.cap='A simple scatter plot.', tidy=FALSE}
data(mtcars) 
plot(mtcars$wt, mtcars$mpg,
     col = "green",  
     main = "MPG as a Function of Weight", 
     xlab = "Weight",
     ylab = "MPG") 
```

**Histogram**
```{r, fig.cap='A simple histogram.'}
hist(mtcars$hp, 
     main = "HP of cars",
     xlab = "hp",
     col = "pink")
```


\newpage
# Appendix 1: R code
### Problem 4
RStudio version: R 4.0.2
```{r Appendix, echo=TRUE, eval=FALSE}
data(mtcars) # loading the whole data set for mtcars
plot(mtcars$wt, mtcars$mpg,
     col = "green",  # color of the points
     main = "MPG as a Function of Weight", # main title of the plot
     xlab = "Weight",
     ylab = "MPG") #a simple scatter plot

hist(mtcars$hp, 
     main = "HP of cars", # main title of the histogram
     xlab = "hp",
     col = "pink")  # A simple histogram
```