---
title: "HW2_wliu3"
author: "Wei Liu"
date: "9/13/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include=FALSE, eval=TRUE, cache=TRUE)
library(data.table)
library(tidyverse)
```


# Problem 1
Finished in the RStudio Cloud

# Problem 2
Finished in the RStudio Cloud

# Problem 3
I think the version control is an very efficient tool to keep track all changes 
and scripts we made, so we can check back anytime and anywhere. This is a good 
way to get the reproducible results.   

# Problem 4

## a. Sensory data from five operators
we are looking at the sensory data from five operators.
http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/Sensory.dat

First, we will get the data from the link above:

```{r sensory_import, cache=FALSE, echo=T}
## getting "http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/Sensory.dat"
url <- "http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/Sensory.dat"
sensory_data_raw<-as.data.frame(fread(url, header=F, fill=T,stringsAsFactors = F, skip=2))
saveRDS(sensory_data_raw,"sensory_data_raw")
sensory_data_raw<-readRDS("sensory_data_raw")
```
Need to tidy the data, some rows are misplaced. There should be variables for each column: items, observation numbers, operators and data. However, five different operators are now in different columns. In addition, observation numbers are missing.


### Method 1: base R 

```{r sensory_tidy_baseR, echo=T, cache=FALSE, include=T}
# check the column and row numbers of sensory_data_raw
dim(sensory_data_raw)
# dim=(30,6)

# Use loop to move the NA in the column V6 to the first column V1
for (i in 1:30) {
  if(is.na(sensory_data_raw$V6[i])) {
  sensory_data_raw[i,-1] <- sensory_data_raw[i,1:5]
  sensory_data_raw[i,1] <- NA
  }
}

# There are 10 items in total, so need fix the items numbers in the first column.
sensory_data_raw$V1 <- rep(1:10, each = 3)

# create a new column V7 for observation numbers.
sensory_data_raw$V7 <- rep(c("ob.1","ob.2","ob.3"),10)
sensory_data_raw1<-sensory_data_raw

# set the column names
colnames(sensory_data_raw1)<-
  c("items", "operator1", "operator2", "operator3", "operator4", "operator5","observations")

# choose the columns of operator1,2,3,4,5 and then combine them into one column. 
data<-c(sensory_data_raw1[,2],sensory_data_raw1[,3],sensory_data_raw1[,4],
        sensory_data_raw1[,5],sensory_data_raw1[,6])

# create the new columns for variables of items, observations, and operators.
items<-rep(rep(1:10, each = 3),5)
observations<-rep(rep(c("ob.1","ob.2","ob.3"),10),5)
operator<-rep(c("operator1","operator2","operator3","operator4","operator5"), each=30)

# bind the above columns together with the sensory data
sensory_data_tidy_br<-as.data.frame(cbind(items,operator,observations,data))

# convert the columns of items and data into numeric
sensory_data_tidy_br$items <- as.numeric(as.character(sensory_data_tidy_br$items))
sensory_data_tidy_br$data <- as.numeric(as.character(sensory_data_tidy_br$data))
```
We have converted the dataframes to tidy data frames using the base functions. Here is a summary of the data:
```{r summary_baseR, include=T, echo=F}
knitr::kable(summary(sensory_data_tidy_br))
```
```{r summary_basebR, include=T, echo=F}
boxplot(sensory_data_tidy_br$data~sensory_data_tidy_br$operator,xlab="Operators",ylab="Data")
```

### Method 2: tidyverse R

```{r sensory_tidy_tidyverse, echo=T, include=T, cache=FALSE}
#stack column using tidyverse
sensory_data_tidy_tv <- sensory_data_raw1 %>% 
  gather(key="operator", value="data", operator1:operator5)

# the head of sensory_data_tidy_tv
knitr::kable(head(sensory_data_tidy_tv))
```
We have converted the dataframes to tidy data frames using the tidyverse functions. Here is a summary of the data:
```{r summary_tidyverseR, include=T, echo=F}
knitr::kable(summary(sensory_data_tidy_tv))
```

## b. Gold Medal performance for Olympic Men's Long Jump

we are looking at Gold Medal performance for Olympic Men’s Long Jump, year is coded as 1900=0.
http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/LongJumpData.dat

First, we will get the data from the link above:

```{r longjump_import, cache=FALSE, echo=T}
## getting "http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/LongJumpData.dat"
url1 <- "http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/LongJumpData.dat"
performance_import_raw <-fread(url1, header=F, fill=T,stringsAsFactors = F, skip=1, data.table = F)
saveRDS(performance_import_raw,"performance_import_raw")
performance_import_raw<-readRDS("performance_import_raw")
```

Need to tidy the data, the data for "year" and "longjump" are in several different columns and the "year" should be converted to normal values.


### Method 1: base R 

```{r performance_tidy_baseR, echo=T, cache=FALSE, include=T}
# check the column and row numbers of sensory_data_raw
dim(performance_import_raw)
# dim=(6,8)

# select the columns for the "year"
year<-c(performance_import_raw[,1], performance_import_raw[,3], performance_import_raw[,5], performance_import_raw[,7])

# select the columns for the "longjump"
longjump<-c(performance_import_raw[,2], performance_import_raw[,4], performance_import_raw[,6], performance_import_raw[,8])

# bind the columns for the "year" and "longjump" together
performance_import_raw1<-cbind(year,longjump)

# creat a new column for converted year values
year1<-performance_import_raw1[,1]+1900
performance_data_tidy_br<-cbind(year1, performance_import_raw1)
performance_data_tidy_br1<-performance_data_tidy_br[,-2]

performance_data_tidy_br2<-na.omit(performance_data_tidy_br1)
```

We have converted the dataframes to tidy data frames using the base functions. Here is a summary of the data:
```{r summary_performance_baseR, include=T, echo=F}
knitr::kable(summary(performance_data_tidy_br2))
```

```{r summary_performance_baseR_plot, include=T, echo=F}
plot(year1, longjump, xlab="year",ylab="longjump performance data")
```

### Method 2: tidyverse R

```{r performance_tidy_tidyverse, echo=T, include=T, cache=FALSE}
#select the columns for the "year"
year_select<-performance_import_raw %>% 
                    select(c(V1,V3,V5,V7))

#stack the columns for the "year"
year_output<-stack(year_select) %>% select(values) %>% 
                    `colnames<-`("year")

#select the columns for the "longjump"
longjump_select<-performance_import_raw %>% 
                    select(c(V2,V4,V6,V8)) 
                    
#stack the columns for the "longjump"
longjump_output<-stack(longjump_select) %>% 
            select(values) %>% 
            `colnames<-`("longjump") 

#bind the columns of "year" and "longjump" together
performance_data_tidy_tv<-cbind(year_output,longjump_output)%>% 
                            mutate(year22=year+1900)%>%
                            select(year22,longjump)%>%
                            na.omit()

# the head of performance_data_tidy_tv
knitr::kable(head(performance_data_tidy_tv))
```
We have converted the dataframes to tidy data frames using the tidyverse functions. Here is a summary of the data:
```{r summary_performance_tidyverseR, include=T, echo=F}
knitr::kable(summary(performance_data_tidy_tv))
```

## c. Brain weight (g) and body weight (kg) data

we are looking at Brain weight (g) and body weight (kg) for 62 species.
http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/BrainandBodyWeight.dat

First, we will get the data from the link above:

```{r weight_import, cache=FALSE, echo=T}
## getting "http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/BrainandBodyWeight.dat"
url2 <- "http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/BrainandBodyWeight.dat"
weight_import_raw<-fread(url2, header=F, stringsAsFactors = F, fill=T, skip=1, data.table = F)
saveRDS(weight_import_raw,"weight_import_raw")
weight_import_raw<-readRDS("weight_import_raw")
```

Need to tidy the data, the data for "brain wt" and "body wt" are in several different columns.


### Method 1: base R 

```{r weight_tidy_baseR, echo=T, cache=FALSE, include=T}
# check the column and row numbers of weight_import_raw
dim(weight_import_raw)
# dim=(21,6)

# select the columns for the "body wt"
body_wt<-c(weight_import_raw[,1], weight_import_raw[,3], weight_import_raw[,5])

# select the columns for the "weight wt"
brain_wt<-c(weight_import_raw[,2],weight_import_raw[,4],weight_import_raw[,6])

# bind the columns for the "body wt" and "weight wt" together
weigiht_data_tidy_br<-cbind(body_wt,brain_wt)

weigiht_data_tidy_br<-na.omit(weigiht_data_tidy_br)
```
We have converted the dataframes to tidy data frames using the base functions. Here is a summary of the data:
```{r summary_weight_baseR, include=T, echo=F}
knitr::kable(summary(weigiht_data_tidy_br))
```

```{r summary_weight_baseR_plot, include=T, echo=F}
plot(body_wt, brain_wt, xlab="body wt",ylab="brain wt")
```

### Method 2: tidyverse R

```{r weight_tidy_tidyverse, echo=T, include=T, cache=FALSE}
#select the columns for the "body wt"
body_wt1<- weight_import_raw %>% 
                    select(c(V1,V3,V5))

#stack the columns for the "body wt"
body_wt2<-stack(body_wt1) %>% select(values) %>% 
                    `colnames<-`("body_wt")

#select the columns for the "brain wt"
brain_wt1<- weight_import_raw %>% 
                    select(c(V2,V4,V6)) 

#stack the columns for the "brain wt"
brain_wt2<-stack(brain_wt1) %>% 
            select(values) %>% 
            `colnames<-`("brain_wt") 

weight_data_tidy_tv<-cbind(body_wt2,brain_wt2) %>% na.omit()

# the head of weight_data_tidy_tv
knitr::kable(head(weight_data_tidy_tv))
```
We have converted the dataframes to tidy data frames using the tidyverse functions. Here is a summary of the data:
```{r summary_weight_tidyverseR, include=T, echo=F}
knitr::kable(summary(weight_data_tidy_tv))
```

## d. Triplicate measurements of tomato yield

Triplicate measurements of tomato yield for two varieties of tomatos at three planting densities.
http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/tomato.dat

First, we will get the data from the link above:

```{r tomato_import, cache=FALSE, echo=T}
## getting "http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/tomato.dat"
url3 <- "http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/tomato.dat"
tomato_import_raw<-fread(url3)
saveRDS(tomato_import_raw,"tomato_import_raw")
tomato_import_raw<-readRDS("tomato_import_raw")
```

Need to tidy the data, several different values for "tomato yield" are in one column.


### Method 1: base R 

For the base R method, I am still working on find a way without using "select" and "separate"
```{r tamoto_tidy_baseR, echo=T, cache=FALSE, include=T}
# select the density columns for "10000","20000" and "30000"
tomato_import_raw1<-select(tomato_import_raw, "10000","20000","30000") 

# separate the yield values for "10000","20000" and "30000"
tomato_import_raw2<-separate(tomato_import_raw1, "10000", into = c("10000m1", "10000m2","10000m3"),sep=",")
tomato_import_raw3<-separate(tomato_import_raw2, "20000", into = c("20000m1", "20000m2","20000m3"),sep=",")
tomato_import_raw4<-separate(tomato_import_raw3, "30000", into = c("30000m1", "30000m2","30000m3"),sep=",")

# transpose the data and convert to numeric
tomato_import_raw4<-t(tomato_import_raw4)
tomato_import_raw5<-c(tomato_import_raw4[,1],tomato_import_raw4[,2])
tomato_import_raw6<-as.numeric(tomato_import_raw5)

# create the new columns for tamato_type, density, and observations 
tomato_type<-rep(c("tomato1","tamato2"),each=9)
density<-rep(rep(c(10000,20000,30000),each=3),2)
observations<-rep(rep(c("ob.1","ob.2","ob.3"),3),2)

# bind the above columns into one dataset
tomato_data_tidy_br<-as.data.frame(cbind(tomato_type,density,observations,tomato_import_raw6))
colnames(tomato_data_tidy_br)<-c("tomato_type","density","observations","yield")

tomato_data_tidy_br$yield<-as.numeric(tomato_data_tidy_br$yield)

#the head of tomato_data_tidy_br
knitr::kable(head(tomato_data_tidy_br))
```

We have converted the dataframes to tidy data frames using the base functions. Here is a summary of the data:
```{r summary_tomato_baseR, include=T, echo=F}
knitr::kable(summary(tomato_data_tidy_br))
```

```{r summary_tomato_baseR_plot, include=T, echo=F}
opar<-par(no.readonly = T)
par(opar)

par(mfrow=c(1,2))
plot(density[1:9], tomato_import_raw6[1:9], xlab="density",ylab="tomato1 yield",main = "tomato1", ylim = c(5, 21))
plot(density[9:18], tomato_import_raw6[9:18], xlab="density",ylab="tomato2 yield",main = "tomato2", ylim = c(5, 21))

par(opar)
```

### Method 2: tidyverse R

```{r tamoto_tidy_tidyverse, echo=T, cache=FALSE, include=T}
# select the density columns for "10000","20000" and "30000" and separate the values
tomato_import_raw7 <- tomato_import_raw %>% 
    select("10000","20000","30000") %>%
    separate("10000", into = c("10000m1", "10000m2","10000m3"),sep=",")%>%
    separate("20000", into = c("20000m1", "20000m2","20000m3"),sep=",")%>%
    separate("30000", into = c("30000m1", "30000m2","30000m3"),sep=",")


# transpose the data and convert to numeric
tomato_import_raw7<-t(tomato_import_raw7)
tomato_import_raw8<-c(tomato_import_raw7[,1],tomato_import_raw7[,2])
tomato_import_raw9<-as.numeric(as.character(tomato_import_raw8))


# bind the above columns into one dataset
tomato_data_tidy_tv<-as.data.frame(cbind(tomato_type,density,observations,tomato_import_raw9))

colnames(tomato_data_tidy_tv)<-c("tomato_type","density","observations","yield")
tomato_data_tidy_tv$yield<-as.numeric(tomato_data_tidy_tv$yield)
```
We have converted the dataframes to tidy data frames using the tidyverse function. Here is a summary of the data:
```{r summary_tomato_tidyverse, include=T, echo=F}
knitr::kable(summary(tomato_data_tidy_tv))
```